---
title: "Functionalizing Rasters from NEON"
author: "Jon Weiner"
date: "June 21, 2016"
output: html_document
---

### Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.
1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

### Initialize dataset

Import Libraries

```{r import_libraries}
#import libraries
library(raster)
library(rgdal)
```

Import CHM and set zero values to NA

```{r import_CHM_and_0toNA}
#import TEAK CHM and set zero values to NA
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

chm[chm==0] <- NA
```

### Classify and create histogram

```{r classify_CHM}
#Set up classifications
tree_class <- c(0, 20, 1, 
                20, 40, 3, 
                40, 60, 5)
#Reshape matrix
tree_matrix <- matrix(tree_class,
                ncol = 3,
                byrow = TRUE)
#Classify dataset
classified_trees <- reclassify(chm,
                               tree_matrix)
#Create Histogram of original data (with 0 == NA)
hist(chm,
     main = "TEAK_CHM_Histogram",
     breaks = 20)
abline(v = 2)
abline(v = 6)
abline(v = 56)

plot(classified_trees)
```

### Add legend 

```{r create_legend}

par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(classified_trees,
     col=c("red","blue","green"), # hard code colors, unclassified (0)=white,
		 #N (1) =blue, S(2)=green
     main="North and South Facing Slopes \nLower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Low", "Medium", "High"),  # make sure the order matches the colors, next
       fill = c("red", "blue", "green"),
       bty="n") # turn off border

```




