---
title: "Functionalizing Rasters from NEON"
author: "Jon Weiner"
date: "June 21, 2016"
output: html_document
---

### Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.
1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

### Initialize dataset

Import Libraries

```{r import_libraries}
#import libraries
library(raster)
library(rgdal)
```

Import CHM and set zero values to NA

```{r import_CHM_and_0toNA}
#import TEAK CHM and set zero values to NA
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

chosen_breaks = c(6, 30, 50, 100)

chm[chm==0] <- NA
```

```{r_cleaner_code}


#Function to create a reclassify matrix from a set of breaks
create_height_class_matrix <- function(breaks){
  #get the length of the breaks vector to figure out how many classes we will have
  br.len <- length(breaks)
  
  #initialize height class vector with 0
  height.class.m <- c(0)
  
  #for input of breaks = c(6, 30, 50, 100)
  # we would like something like this
  #c(0, 6, 1, 
  #  6, 30, 2, 
  #  30, 50, 3, 
  #  50, 100, 4)
  for (i in 1:br.len){
    height.class.m <- c(height.class.m, breaks[i-1], breaks[i], i)
  }
  reclass.height.mat <- matrix(height.class.m,
                               ncol = 3, 
                               byrow = TRUE)
}

create_height_class_matrix(chosen_breaks)

#a function to plot density of heights with chosen breaks
#expects a chm raster, a title, and vector of breaks

plot_chm_density <- function(rast.in, title, bins){
  density(rast.in, main = title, xlab = "Height (m)")
  abline(v = bins, col = "red")
}

plot_chm_density(rast.in = chm, 
                 title = "Canopy heights at Teakettle", 
                 bins = chosen_breaks)
```





### Classify and create histogram

```{r classify_CHM}

#Classify dataset
classified_trees <- reclassify(chm,
                               tree_matrix)

par(xpd = FALSE)
#Create Histogram of original data (with 0 == NA)
density(chm,
     main = "TEAK_CHM_Density",
     xlab = "Height (m)")

abline(v = tree_matrix[, 2], col = "red")



#plot(classified_trees)
```

### Add legend 

```{r create_legend}

par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(classified_trees,
     col=c("red","blue","green"), # hard code colors, unclassified (0)=white,
		 #N (1) =blue, S(2)=green
     main="North and South Facing Slopes \nLower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Low", "Medium", "High"),  # make sure the order matches the colors, next
       fill = c("red", "blue", "green"),
       bty="n") # turn off border



pdf(file = "TEAK_class_height.pdf", width = 7, height = 6)




par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(classified_trees,
     col=c("red","blue","green"), # hard code colors, unclassified (0)=white,
		 #N (1) =blue, S(2)=green
     main="North and South Facing Slopes \nLower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Low", "Medium", "High"),  # make sure the order matches the colors, next
       fill = c("red", "blue", "green"),
       bty="n") # turn off border

dev.off()
```

Export to raster

```{r export_raster}

writeRaster(classified_trees,
            filename = "TEAK_class_height_chm.tiff",
            format = "GTiff",
            Naflag = -9999,
            overwrite = TRUE)

sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"),
            "_sessionInfo.txt"))
sessionInfo()
sink()

```




